Refinements for open automata

### 2023-02-27 2021-Refinement/paper/refinementarticleV2.tex
todos and small changes

\begin{document}
%
\title{Refinements for open automata}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{
Rabéa Ameur-Boulifa\inst{1}\orcidID{0000-0002-2471-8012} \and
Quentin Corradi\inst{2}\orcidID{0000-0003-4218-3987} \and
Ludovic Henrio \inst{2}\orcidID{0000-0001-7137-3523} \and
Eric Madelaine \inst{3}\orcidID{0000-0002-5552-5993}}
%
\authorrunning{Rab\'ea Ameur-Boulifa et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{LTCI, T\'el\'ecom Paris, Institut Polytechnique de Paris, France
\email{first.last@telecom-paris.fr}\\
\and
Universit\'e Lyon, EnsL, UCBL, CNRS, Inria,  LIP, France\\
\email{first.last@ens-lyon.fr}
 \and
INRIA Sophia Antipolis M\'edit\'erann\'ee, UCA,  France
\email{first.last@inria.fr}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
%\TODO{ The abstract should briefly summarize the contents of the paper in 150--250 words.}

@@

@@

\begin{example} As a running example, we consider a data base example inspired from \cite{Gorrieri:2001}. The open automaton \texttt{database}  modelling the behaviour of data base is depicted in Fig. \ref{Fig:SpecOA}. The \texttt{database} can be queried using an operation \texttt{qry} and updated using an operation \texttt{upd}. The associated automaton has a single hole:  a counter ($\texttt{cnt}$)  depicting  the behaviour of the timer. When the \texttt{database}  performs a query,  it sets the new time limit of the counter component and the exposed action to the environment is an observable  action \texttt{qry}.
The \texttt{database} updates and  switches when it agrees with the counter component the time is over, and the exposed action is \texttt{upd}.
The open automaton \texttt{counter}    depicting  the behaviour of the timer is illustrated in Fig. \ref{Fig:counter}. 
\end{example}
\begin{figure}[h]
 \centering
   \includegraphics[width=.9\textwidth]{Figures/database.pdf}
   \caption{Open Automaton. It exposes an unspecified counter in the "cnt" hole
 to count external tick actions. The system sets a counter when a query is enabled and updates  when  the time is over.\label{Fig:SpecOA}}
\end{figure}

@@

We define now a simulation between OA that have the same set of holes. This relation thus focuses on the simulation of transitions and uses the previous definition to prevent the refined automaton from non-deterministically reaching states that introduce deadlocks in a scenario planned without deadlock in the specification.
\TODO{Changement de la definition de simulation pour selon la simulation de Olga}

@@

\TODO{To discuss this definition}

@@

\begin{definition}[Deadlock reduction]\label{def:dpwd}\\
Let \(\OAg[1]\) and \(\OAg[2]\) be two OAs.
A relation on OA configurations \(R: S_1 \times S_2 \to \rformulas[V_1 \uplus V_2]\) is deadlock reducing if  it satisfies the following\footnote{Note that variables of $T_1$ are existentially quantified in the proposition.}:
\begin{multline*}
 \forall \mpar{s_1, s_2} \in S_1 \times S_2,\\ V_1 \uplus V_2 \uplus \biguplus_\subbox{t_2 \in \fOT{s_2}} \fvars{t_2} \vdash \left(R\mpar{s_1, s_2} \wedge \bigvee_\subbox{t_2 \in \fOT{s_2}} \fguard{t_2} \implies \bigvee_\subbox{t_1 \in \fOT{s_1}} \fguard{t_1} \right)
\end{multline*}
\end{definition}

@@

\TODO{Fin reintroduction}

@@

\begin{theorem}[Context refinement]\label{thm:ContextRefinement}
Let $A_1$, $A_2$ and $A_3$ be three open automata with $\wrel{A_1}{A_2}{H}$. 
Let $J_3$ be the set of holes of $A_3$ and suppose that $k \in H$.
%Suppose that \(k \in H\) and that \(A_1\subst{A_3}{k}\) is non-blocking.
We have: \[\wrel{A_1\subst{A_3}{k}}{A_2\subst{A_3}{k}}{J_3 \uplus H \setminus \mbrc{k}}\]
\end{theorem}
\RAB{A t-on des conditions sur H?}

@@

\begin{theorem}[Congruence]
Let $A_1$, $A_2$ and $A_3$ be three open automata with $\wrel{A_2}{A_3}{H}$. 
 $\wrel{A_1}{A_2}{H}$.  Let $J_3$ be the set of holes of $A_3$ and suppose that \(k \in H\).
We have: \[\wrel{A_1\subst{A_2}{k}}{A_1\subst{A_3}{k}}{J_1 \uplus H \setminus \mbrc{k}}\]
\end{theorem}

\begin{example} \RAB{cet exemple illustre plutot  que la composition est un raffinement, i.e.,  $\wrel{A_1\subst{A_2}{k}}{A_1}{J_1   \setminus \mbrc{k}}$ doit-on ajouter ce resultat comme un lemme?}

>>>>

\begin{document}
%
\title{Refinements for open automata}

%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{
Rabéa Ameur-Boulifa\inst{1}\orcidID{0000-0002-2471-8012} \and
Quentin Corradi\inst{2}\orcidID{0000-0003-4218-3987} \and
Ludovic Henrio \inst{2}\orcidID{0000-0001-7137-3523} \and
Eric Madelaine \inst{3}\orcidID{0000-0002-5552-5993}}
%
\authorrunning{Rab\'ea Ameur-Boulifa et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{LTCI, T\'el\'ecom Paris, Institut Polytechnique de Paris, France
\email{first.last@telecom-paris.fr}\\
\and
Universit\'e Lyon, EnsL, UCBL, CNRS, Inria,  LIP, France\\
\email{first.last@ens-lyon.fr}
 \and
INRIA Sophia Antipolis M\'edit\'erann\'ee, UCA,  France
\email{first.last@inria.fr}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
%\TODO{ The abstract should briefly summarize the contents of the paper in 150--250 words.}

@@

\TODO{TODO list}

* case 2 of the proofs for both theorems

* example

* positioning

* global rewriting

@@

\begin{example} As a running example, we consider a data base example inspired from \cite{Gorrieri:2001}. The open automaton \texttt{database}  modelling the behaviour of data base is depicted in Fig. \ref{Fig:SpecOA}. The \texttt{database} can be queried using an operation \texttt{qry} and updated using an operation \texttt{upd}. The associated automaton has a single hole:  a counter ($\texttt{cnt}$)  depicting  the behaviour of the timer. When the \texttt{database}  performs a query,  it sets the new time limit of the counter component and the exposed action to the environment is an observable  action \texttt{qry}.
The \texttt{database} updates and  switches when it agrees with the counter component the time is over, and the exposed action is \texttt{upd}.
The open automaton \texttt{counter}    depicting  the behaviour of the timer is illustrated in Fig. \ref{Fig:counter}. 
\end{example}
\begin{figure}[h]
 \centering
   \includegraphics[width=.9\textwidth]{Figures/database.pdf}
   \caption{Open Automaton. It exposes an unspecified counter in the "cnt" hole
 to count external tick actions. The system sets a counter when a query is enabled and updates  when  the time is over.\label{Fig:SpecOA}} \TODO{refaire l exemple pour l instant il a pas vraiment de sens}
\end{figure}

@@

We define now a simulation between OA that have the same set of holes. This relation thus focuses on the simulation of transitions and uses the previous definition to prevent the refined automaton from non-deterministically reaching states that introduce deadlocks in a scenario planned without deadlock in the specification.

@@

@@

\TODO{remove this def}
\begin{definition}[Deadlock reduction]\label{def:dpwd}\\
Let \(\OAg[1]\) and \(\OAg[2]\) be two OAs.
A relation on OA configurations \(R: S_1 \times S_2 \to \rformulas[V_1 \uplus V_2]\) is deadlock reducing if  it satisfies the following\footnote{Note that variables of $T_1$ are existentially quantified in the proposition.}:
\begin{multline*}
 \forall \mpar{s_1, s_2} \in S_1 \times S_2,\\ V_1 \uplus V_2 \uplus \biguplus_\subbox{t_2 \in \fOT{s_2}} \fvars{t_2} \vdash \left(R\mpar{s_1, s_2} \wedge \bigvee_\subbox{t_2 \in \fOT{s_2}} \fguard{t_2} \implies \bigvee_\subbox{t_1 \in \fOT{s_1}} \fguard{t_1} \right)
\end{multline*}
\end{definition}

@@

\subsection{Automaton refinement}

@@

\begin{theorem}[Context refinement]\label{thm:ContextRefinement}
Let $A_1$, $A_2$ and $A_3$ be three open automata with $\wrel{A_1}{A_2}{H}$. 
Let $J_3$ be the set of holes of $A_3$ and suppose that $k \in H$.
Suppose that \(k \in H\) and that \(A_1\subst{A_3}{k}\) is non-blocking.
We have: \[\wrel{A_1\subst{A_3}{k}}{A_2\subst{A_3}{k}}{J_3 \uplus H \setminus \mbrc{k}}\]
\end{theorem}

@@

\begin{theorem}[Congruence]
Let $A_1$, $A_2$ and $A_3$ be three open automata with $\wrel{A_2}{A_3}{H}$. 
  Let $J_3$ be the set of holes of $A_3$ and suppose that \(k \in H\). Suppose additionally that the composition $A_1\subst{A_2}{k}$ is non-blocking.
We have: \[\wrel{A_1\subst{A_2}{k}}{A_1\subst{A_3}{k}}{J_1 \uplus H \setminus \mbrc{k}}\]
\end{theorem}

\begin{example}
